Mixed effect tweedie model in TMB

### 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
batch cross-validation of temp models

```{r setup}
library(tidyverse)
library(sdmTMB)

d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))

#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
set.seed(392011149)
d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl$depth2 <- (d_trawl$depth_scaled)^2
d_trawl$depth3 <- (d_trawl$depth_scaled)^3
```

@@

nrow(d_fit) # 3774
nrow(d_withheld) # 943
unique(d_withheld$year) 
unique(d_fit$year)
length(unique(d_fit$year))

# FIXME: need to work out a bug in production on this new data set
# It's crashing TMB.

#d_fit <- d_trawl
#d_withheld <- d_trawl
```

```{r}
spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
plot_spde(spde)
```

Functions for model exploration:
```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", colour = column)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed()
}

Various candidate models:

@@

```{r, eval=FALSE}
m_temp2 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

@@

# Extrapolate:

#FIXME: this is crashing R

```{r}
m_temp2_rw_poly <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled,2),
  include_spatial = FALSE, ar1_fields = FALSE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)
```

```{r}
predictions_nd <- predict(m_temp2_rw_poly, newdata = qcs_grid)
#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
```

# A short function for plotting our predictions:

```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

>>>>

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)

d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
set.seed(392011149)
#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
```

### Functions for model exploration:
```{r}
generate_cv_args <- function(
                formula = formula,
                time_varying = time_varying,
                ar1_fields = c(FALSE,TRUE), 
                include_spatial = c(FALSE,TRUE),
                n_knots = c(30,80) ){
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots
  )
  args <- as_tibble(arguments)
  args
}

plot_cv_meshes <- function(cv_out) {
  k <- max(cv_out$data$cv_fold)
  op <- graphics::par(
      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
    )
  for (i in seq_len(k)){
    plot_spde(cv_out$models[[i]]$spde)
  }
graphics::par(op)
}

# 10-fold cross-validation

## All fixed slope models
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
    time_varying = list(NULL),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE,TRUE),
    n_knots = c(30,80,150)
  )

plan(multiprocess)
sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
cv_args1 <- cv_args
cv_loglik <- c()
for (i in seq_len(nrow(cv_args1))) {
  cv_loglik[[i]] <- round(sdmTMB_cv_batch1[[i]]$sum_loglik, 1)
}
cv_args1$formula <- as.character(cv_args1$formula)
cv_args1$cv_loglik <- cv_loglik
View(cv_args1)
cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
glimpse(cv_args_notv)
write.csv(cv_args_notv, file="cv-all-non-time-varying.csv")
saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
# readRDS(file = "sdmTMB_cv_nontv.rds")
```

## All time-varying slope models
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled,
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE,TRUE),
    n_knots = c(30,80,150)
    )

plan(multiprocess)
sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)

cv_loglik <- c()
for (i in seq_len(nrow(cv_args))) {
  cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
}
cv_args2 <- cv_args
cv_args2$formula <- as.character(cv_args$formula)
cv_args2$time_varying <- as.character(cv_args$time_varying)
cv_args2$cv_loglik <- cv_loglik
glimpse(cv_args2)
cv_args_tv <- as.data.frame(cv_args2) 
write.csv(cv_args_tv, file="cv-all-time-varying.csv")
saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
# readRDS(file = "sdmTMB_cv_nontv.rds")
```

# A functions for plotting our predictions:

```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", colour = column)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

## Plotting predictions

```{r}
predictions_nd <- predict(m_temp2_rw_nosp, newdata = qcs_grid)
#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
```

########################
########################

Various candidate models

########################
########################

```{r}
spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
p <- plot_spde(spde)
p
```

@@

nrow(d_fit) # 3774
nrow(d_withheld) # 943
unique(d_withheld$year) 
unique(d_fit$year)
length(unique(d_fit$year))
```

Misc models
```{r}
temp2 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALSE, include_spatial = FALSE)
plot_cv_meshes(temp2)
sum(temp2$sum_loglik)

temp3 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALSE, include_spatial = FALSE)
sum(temp3$sum_loglik)

# allow slope with depth to vary through between years with a random walk
temp1_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALSE, include_spatial = FALSE)
sum(temp1_rw$sum_loglik)

temp2_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALSE, include_spatial = FALSE)
sum(temp2_rw$sum_loglik)

temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALSE, include_spatial = FALSE)
sum(temp3_rw$sum_loglik)
```

@@

```{r, eval=FALSE}
#d_trawl$depth2 <- (d_trawl$depth_scaled)^2
#d_trawl$depth3 <- (d_trawl$depth_scaled)^3

m_temp2 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

@@

# there is probably a problem with doing this... but need to figure out how to add poly args to predict function
```{r eval=FALSE}
m_temp2_rw_poly <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled,2),
  include_spatial = FALSE, ar1_fields = FALSE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)
```
